---
title: "Snap Tool"
author: "Emma Jones"
date: "January 2, 2019"
output: html_document
---

```{r setup, include=FALSE}

# Run in R3.5.1

knitr::opts_chunk$set(echo = TRUE)
#library(dplyr)
#library(readr)
#library(purrr)
library(tidyverse)
library(sf)
```

This script demonstrates an initial workflow to retrieve attribute information from NHD flowlines. All analysis will comply with tidy data workflows. 

### Initial Thoughts on Snapping Points to Lines

**Background** 
I built an S3 version of this workflow in ~2015 to snap VDEQ monitoring locations to a NHD polyline shapefile with embedded Water Quality Standards information. This process was an initial step in automating WQS assessment workflows throughout the state. After a few years of project stagnation, renewed interest in automating assessments came about in 2017 with increased shiny app acceptance within the agency. The goal of this snapping project is to develop a function that can take point information and associate the nearest line feature for extraction of attribute info. 

**Why a new function?** 
After thorough exploration of native sf snapping functions (sf::st_snap, sf::st_nearest_feature) I have determined that they do not fully meet needs for flow line applications. These functions only return the closest feature to an input point, which can mask errors in snapping to the 'correct' flow line. Though a number of lat/lng fall close enough to stream segments to use the above mentioned snapping functions, there are cases where a point falls close to 2+ flow lines. These require manual QA to identify where the point intended to snap to. By only returning the closest feature, these cases are masked from users and can potentially snap to incorrect stream segments. There is value in knowing how many segments are identified within a given buffer distance.

### Snapping Functions that output Points

These nested functions accomplish the task of taking input spreadsheet point data and snapping to nearest line segments. **This version of the functions output point(s) instead of a list of lines and points.** See snapFunctions.Rmd for previous iterations of snapping function.

Bring in test dataset. I clipped a small bit of the WQS layer from New river basin for test NHD. The points are a few real and a few made up (I picked sites that would purposefully grab too many stream geometries) for testing. I made sure to start with a spreadsheet form to make sure the function will be most useful to people. The sf transformation steps can happen in an outer function and can build out something that tests incoming dataset and adjusts workflow accordingly.

```{r newTestData}
WQS <- st_read('data/WQS2018_BRRO_albers_mini.shp')

probSites_xl <- read_csv('data/probSites_mini.csv')

probSites_sf <- st_as_sf(probSites_xl, 
                    coords = c("LongitudeD", "LatitudeDD"), # for point data
                    remove = F, # don't remove these lat/lon cols from df
                    crs = 4269) %>% # add projection, needs to be geographic for now bc entering lat/lng, 
  st_transform( st_crs(WQS))# project to Albers equal area for snapping

# make sure both layers have same CRS
identical(st_crs(WQS),st_crs(probSites_sf))
```


Starting small, build first function that just buffers single site X distance and lets user know if there is a match with input polyline file. Still using my version of inner function because I don't want to test to see if a buffer intersects a stream, then re-grab that stream's unique identifier in a separate step. I think it makes more sense to just do the st_intersect once even though the apply() functionality is a tad faster, I would need to st_intersect twice to get out the info I need.


```{r bufferFunction}
# For testing
POINT <- probSites_sf[1,]
MULTILINESTRING <- WQS
distance <- 40

snap_bufferMethod <- function(POINT, MULTILINESTRING, distance){
  step1 <- st_buffer(POINT,dist = distance)
  st_zm(MULTILINESTRING) %>% 
    filter(st_intersects(., st_zm(step1), sparse = FALSE))
}

snap_bufferMethod(probSites_sf[1,],WQS,5)
snap_bufferMethod(probSites_sf[1,],WQS,40)
```

Then build outer function that accepts sequence of buffer distances for a single site.

```{r buffer series}
# For testing
POINT <- probSites_sf[4,]#probSites_sf[1,]#probSites_sf[5,]#
MULTILINESTRING_UID_colname <- "OBJECTID"
MULTILINESTRING <- WQS
bufferDistances <- seq(10,50,by=10)

snap_Point_to_Feature <- function(POINT, # sf POINT file
                                  MULTILINESTRING, # stream network
                                  MULTILINESTRING_UID_colname, # as.character(name of unique identifier in POINT file)
                                  bufferDistances # numeric sequence of distances to run buffer, these will be in
                                  ){              # the unit of the POINT and MULTILINESTRING files
  
   x <- 0
  repeat {
   x <- x + 1
   b <- snap_bufferMethod(POINT,MULTILINESTRING,bufferDistances[x])
   if (nrow(b) > 0 | x == length(bufferDistances)) break   }
  

  if( nrow(b) == 0 ){
    c <- mutate(POINT, `MULTILINESTRING Unique Identifier` = NA,
                `Buffer Distance` = paste('No connections within', max(bufferDistances),
                                        st_crs(POINT)$units, sep = ' '))
  } else {
       cn <- as.character(st_set_geometry(b,NULL) %>% select_(MULTILINESTRING_UID_colname) %>% flatten_chr())
       c <- POINT %>%
         slice(rep(1:n(), each = length(cn))) %>% # replicate rows of POINT file length(cn) times to mutate cn info
         mutate(`MULTILINESTRING Unique Identifier` = cn,
                   `Buffer Distance` = paste(bufferDistances[x], 
                                             st_crs(POINT)$units, sep = ' ')) }
  return(c)
}

snap_Point_to_Feature(probSites_sf[1,], WQS, 'OBJECTID',  seq(10,50, 5))
``` 

But if you want to make a dataset with sites that did connect to something and sites that didn't it causes problems in MULTILINESTRING object.

```{r problemsIdentified}
z1 <- snap_Point_to_Feature(probSites_sf[1,], WQS, 'OBJECTID', seq(10,50,by=10)) # one connection
z3 <- snap_Point_to_Feature(probSites_sf[4,], WQS, 'OBJECTID', seq(10,50,by=10)) # three connections
z0 <- snap_Point_to_Feature(probSites_sf[5,], WQS, 'OBJECTID', seq(10,50,by=10)) # no connection

# Ideally we would combine them now to output to a user, but can't bc sf issues with bind_rows

z4 <- suppressWarnings(bind_rows(z1,z3,z0) )
z5 <- suppressWarnings(rbind(z1,z3,z0, deparse.level = 1)) 
```


Now need to extend this to MULTIPOINT features so user can easily use with spreadsheet-like input. This function outputs a list with one sf object for things that connected to segment within buffer and one smaller tbl that identifies sites that need attention. I didn't love this initally, but I think it offers a somewhat elegant solution to the problem because a single sf output with sites that worked and ones that didn't would still need to be filtered to get rid of empty geometry rows before manipulating further.

Also, don't love the loop in the function, costly. Maybe a Jenny Bryan list solution could speed up processing here? leaving it at loop for now.


```{r multipoint with list output}

snap_Points_to_Feature_List <- function(MULTIPOINT, # sf MULTIPOINT file
                                  POINT_UID_colname, # as.character(name of unique identifier in POINT file)
                                  MULTILINESTRING, # stream network
                                  bufferDistances # numeric sequence of distances to run buffer, these will be in
                                  ){              # the unit of the MULTIPOINT and MULTILINESTRING files)
  # Make a list to store the two types of results
  out_list <- list(sf_output = list(), tbl_output = list() )
  
  # Don't love using a loop here but can't figure out better way at present
  for(i in 1:nrow(MULTIPOINT)) {
    print(paste('Snapping Point ',i,' of ',nrow(MULTIPOINT), sep = ''))
    z <- snap_Point_to_Feature(MULTIPOINT[i,], POINT_UID_colname, MULTILINESTRING, bufferDistances)
    if("sf" %in% class(z)){
      if( length(out_list$sf_output) == 0 ){
        out_list$sf_output <- z
      } else { 
        out_list$sf_output <- suppressWarnings( out_list$sf_output <- rbind(out_list$sf_output,z) ) }
    } else {
      out_list$tbl_output <-  suppressWarnings( out_list$tbl_output %>% bind_rows(z) )
    }
  }
  

    
  
  
  # Report Results
  print('Use objectName$sf_output to view successful connections.')
  print('Use objectName$tbl_output to view sites that did not connect to any segments within the input buffer distances.')
  
  return(out_list)
}

```



```{r multipoint with list output}

snap_Points_to_Feature_List2 <- function(MULTIPOINT, # sf MULTIPOINT file
                                  POINT_UID_colname, # as.character(name of unique identifier in POINT file)
                                  MULTILINESTRING, # stream network
                                  bufferDistances # distance to run buffer, this will be in
                                  ){              # the units of the MULTIPOINT and MULTILINESTRING files)
  # Make a list to store the two types of results
  out_list <- list(sf_output = list(), tbl_output = list() )
  
  # Don't love using a loop here but can't figure out better way at present
  for(i in 1:nrow(MULTIPOINT)) {
    print(paste('Snapping Point ',i,' of ',nrow(MULTIPOINT), sep = ''))
    z <- snap_Point_to_Feature(MULTIPOINT[i,], POINT_UID_colname, MULTILINESTRING, bufferDistances)
    if("sf" %in% class(z)){
      if (nrow(z) > 1){
        y <- z %>%
          dplyr::select(`Point Unique Identifier`,`Buffer Distance`)%>%
          dplyr::slice(1) %>%
          st_set_geometry(NULL)
        out_list$tbl_output <- tibble(`Point Unique Identifier` = y$`Point Unique Identifier`,
                    `Buffer Distance` = paste0('Multiple Connections within ',y$`Buffer Distance`))
        
      }
      if (length(out_list$sf_output) == 0 ){
        out_list$sf_output <- z
      } else { 
        out_list$sf_output <- suppressWarnings( out_list$sf_output <- rbind(out_list$sf_output,z) ) }
    } else {
      out_list$tbl_output <-  suppressWarnings( out_list$tbl_output %>% bind_rows(z) )
    }
  }

    
  
  # Report Results
  print('Use objectName$sf_output to view successful connections.')
  print('Use objectName$tbl_output to view sites that did not connect to any segments within the input buffer distances.')
  
  return(out_list)
}

```

